Higgs Assisted Razor Search for Higgsinos at a 100 TeV $pp$ Collider

### 2019-12-17 main.tex
Shufang's edits

 % Why go beyond the MSSM?
Though a natural MSSM spectrum has the potential to tame the hierarchy problem,
it is under siege from recent data from the Large Hadron Collider (LHC)
\cite{Aaboud:2018ujj, Sirunyan:2018vjp}.  A compelling alternative scenario
comes in the form of split supersymmetry (split SUSY) \cite{Wells:2003tf,
ArkaniHamed:2004yi, Giudice:2004tc}. In this scenario, the lightest
superpartners are the fermionic ones, on the scale of 1-10 TeV, while the
scalar superpartners can be much heavier, on the scale of 100 - 1000 TeV. In
exchange for accepting some level of fine-tuning, we obtain numerous benefits,
including the suppression of flavor-changing neutral currents and greater
compatibility with data from CP-violation experiments.  The low lying fermionic
superpartners consist of neutralinos and charginos (collectively known as
electroweakinos), and gluinos. The current LHC search limits on gluinos are
already around 2 TeV for prompt decay and 2.5 TeV for disappearing track
searches~\cite{CMS-PAS-SUS-19-005}. The limits on electroweakinos with
  heavy scalar superparticles, however, are still relatively weak: between 150
  GeV $-$ 1.0 TeV, depending on the neutralino/chargino species, decay modes,
  and their mass splittings~\cite{ATL-PHYS-PUB-2019-022, Sirunyan:2017lae,
    Sirunyan:2018ubx, Aaboud:2018htj, Aaboud:2018zeb, Aad:2019qnd, Aad:2019vvf,
  Aad:2019vnb, Aaboud:2018ngk, Aaboud:2018sua, Aaboud:2018jiw, Aaboud:2017leg,
Aaboud:2017mpt, Aaboud:2017nhr}.  In this paper, we will focus on the discovery
potential of the neutralinos and charginos at a 100 TeV $pp$ collider. {\bf
Can not find Ref. 9}

@@

% Why a 100 TeV collider?
A 100 TeV $pp$ collider would open up an immense number of physics
opportunities not afforded to the 14 TeV LHC. \st{including exploring the most
interesting regions of the split SUSY parameter space}~\cite{Arkani-Hamed:2015vfh}.
% Status of 100 TeV planning
Such a collider would be a natural next step after the LHC, and is being
actively discussed in the particle physics community, with two major proposals
being the Future Circular Collider (FCC)-hh by CERN \cite{FCC-hh}, and the
Super Proton Proton Collider (SppC)  in China \cite{CEPC}.  As the next energy
frontier machine, it is crucial to fully explore the physics
potential of a 100 TeV $pp$ collider, especially for new particles that are
either too heavy or too rare to be produced at the
LHC~\cite{Arkani-Hamed:2015vfh,Contino:2016spe,Golling:2016gvc,Mangano:2016jyj}.
In particular, studies have already been performed on the prospects of
discovering neutralino dark matter for compressed \cite{Low:2014cba,Cirelli:2014dsa,Mahbubani:2017gjh,Mahbubani:2018tin,Han:2018wus,Bramante:2014tba, Berlin:2015aba, 
Fukuda:2017jmk, Saito:2019rtg} and
well-separated \cite{Gori:2014oua,Acharya:2014pua,diCortona:2014yua} neutralino spectra.

@@

We simulated parton-level events using \texttt{MadGraph5 v2.3.2.2} and
\texttt{MadEvent} \cite{Alwall:2014hca}, then passed those events to Pythia 6
\cite{Sjostrand:2006za} for showering and hadronization. Finally, we used
\texttt{Delphes 3} \cite{deFavereau:2013fsa} to perform a fast, parametrized
detector simulation, with the detector card devised by the FCC-hh working
group\footnote{\url{https://github.com/HEP-FCC/FCCSW/blob/master/Sim/SimDelphesInterface/data/FCChh_DelphesCard_Baseline_v01.tcl}}.
For the backgrounds, we allowed up to one additional jet in the final state, to
approximate NLO QCD effects, and performed MLM matching with the \texttt{xqcut}
parameter set to 40 GeV. The Higgsino pair production cross sections were
calculated using \texttt{Prospino2}~\cite{Beenakker:1999xh} at NLO.  
To decouple the Wino, we set its mass parameter $M_2$ to 3 TeV.

@@

Table \ref{tab:cc_cutflowtable} shows the cut efficiencies for a representative
signal benchmark point with $\mu =$ 1 TeV and $M_1$ = 25 GeV.   The
cross-sections of the backgrounds are obtained from MadEvent after performing
the MLM matching procedure. Their small size reflects the $m_{\ell\ell}$ and
$\slashed{E}_T$ cuts we imposed at the parton level. For all the benchmark
points, we require a minimum of three background events (at the MC level, i.e.
unweighted) left over after cuts\footnote{\ap{This is inspired by the upper
    one-sided limit for the mean of a Poisson variable given zero observed
    events, for a confidence level of 95\% (see Table 39.3 of
    \cite{PhysRevD.98.030001}), however, it does not reflect it exactly (since
    we required a minimum on unweighted rather than weighted events).  We
    observed that as we generated more unweighted background events, the
    discovery reach of our analysis showed a steady increase. This, combined
    with the fact that we were pushing the limits of our storage and computing
    capacity (we generated $\approx 10$ million $tt$ events, $\approx 26$
    million $tbW$  events and $100$K $bbWW$ events - all unweighted), lead us
    to believe that the projected reaches presented in this paper are
conservative ones. The ratio of the cross section of the generated events to
that expected at a 100 TeV $pp$ collider with 3 ab$^{-1}$ is $\approx 0.2$ for
$tt$, 0.05 for $tbW$, $0.46$ for $bbWW$}}. We can see that, after applying our
cuts, $tbW$ remains as the dominant backgrounds, followed by $tt$.  The values
of the razor variable cuts were chosen to maximize the significance,
$S/\sqrt{B}$ (calculated for an integrated luminosity of 3000 fb$^{-1}$), shown
in the last column.

@@

For each signal mass combination and each background process, we preselected
events that passed the lepton trigger, contained two SFOS leptons and two
$b$-tagged jets, and calculated a number of kinematic variables for each event.   We then
placed these kinematic variables in an array, with each row corresponding to an
event, and the columns corresponding to the kinematic variables. A mixture of
low-level and high-level kinematic variables was shown to have the greatest
effectiveness. The kinematic variables chosen were $m_{\ell\ell}$, $m_{bb}$,
$M_R$, $M_T^R$, $\slashed{E}_T,$ the total hadronic transverse energy $H_T$,
and the transverse momenta of individual final state leptons and $b$-quarks:
$p_T^{\ell_1}$, $p_T^{\ell_2}$, $p_T^{b_1}$, and $p_T^{b_2}$. We then divided
the generated events into training and test sets, with the training set
comprising 75\% of signal events and 30\% of background events.  \Shufang{More details here -- Referee comment 9.} We used the
\texttt{scikit-learn} package  \cite{Pedregosa2011} to implement our analysis.
A boosted decision tree classifier was trained with 1000 weak learners and a
learning rate of 0.025.

@@

\st{For certain values of the
cuts on kinematic variables or the score of the classifier, no events survived
for one or more of the background components. To guard against overly optimistic
estimates of the significance $S/\sqrt{B}$, we set a lower bound of 3 for the
number of Monte Carlo (MC) events after cuts for each background component, corresponding to
the 95\% confidence interval for a Poisson distribution.}\Shufang{SS: Is it 3
MC events, or 3 weighted events after CS values used? Referee comment 10.}
\Adarsh{I added a footnote about this - see footnote 5. I think we can remove
this paragraph...}

>>>>

 % Why go beyond the MSSM?
Though a natural MSSM spectrum has the potential to tame the hierarchy problem,
it is under siege from recent data from the Large Hadron Collider (LHC)
\cite{Aaboud:2018ujj, Sirunyan:2018vjp}.  A compelling alternative scenario
comes in the form of split supersymmetry (split SUSY) \cite{Wells:2003tf,
ArkaniHamed:2004yi, Giudice:2004tc}. In this scenario, the lightest
superpartners are the fermionic ones, on the scale of 1-10 TeV, while the
scalar superpartners can be much heavier, on the scale of 100 - 1000 TeV. In
exchange for accepting some level of fine-tuning, we obtain numerous benefits,
including the suppression of flavor-changing neutral currents and greater
compatibility with data from CP-violation experiments.  The low lying fermionic
superpartners consist of neutralinos and charginos (collectively known as
electroweakinos), and gluinos. The current LHC search limits on gluinos are
already around 2 TeV for prompt decay and 2.5 TeV for disappearing track
searches~\cite{CMS-PAS-SUS-19-005}. The limits on electroweakinos with
  heavy scalar superparticles, however, are still relatively weak: between 150
  GeV $-$ 1.0 TeV, depending on the neutralino/chargino species, decay modes,
  and their mass splittings~\cite{ATL-PHYS-PUB-2019-022, Sirunyan:2017lae,
    Sirunyan:2018ubx, Aaboud:2018htj, Aaboud:2018zeb, Aad:2019qnd, Aad:2019vvf,
  Aad:2019vnb, Aaboud:2018ngk, Aaboud:2018sua, Aaboud:2018jiw, Aaboud:2017leg,
Aaboud:2017mpt, Aaboud:2017nhr}.  In this paper, we will focus on the discovery
potential of the neutralinos and charginos at a 100 TeV $pp$ collider.

@@

% Why a 100 TeV collider?
A 100 TeV $pp$ collider would open up an immense number of physics
opportunities not afforded to the 14 TeV LHC~\cite{Arkani-Hamed:2015vfh}.
% Status of 100 TeV planning
Such a collider would be a natural next step after the LHC, and is being
actively discussed in the particle physics community, with two major proposals
being the Future Circular Collider (FCC)-hh by CERN \cite{FCC-hh}, and the
Super Proton Proton Collider (SppC)  in China \cite{CEPC}.  As the next energy
frontier machine, it is crucial to fully explore the physics
potential of a 100 TeV $pp$ collider, especially for new particles that are
either too heavy or too rare to be produced at the
LHC~\cite{Arkani-Hamed:2015vfh,Contino:2016spe,Golling:2016gvc,Mangano:2016jyj}.
In particular, studies have already been performed on the prospects of
discovering neutralino dark matter for compressed \cite{Low:2014cba,Cirelli:2014dsa,Mahbubani:2017gjh,Mahbubani:2018tin,Han:2018wus,Bramante:2014tba, Berlin:2015aba, 
Fukuda:2017jmk, Saito:2019rtg} and
well-separated \cite{Gori:2014oua,Acharya:2014pua,diCortona:2014yua} neutralino spectra.

@@

We simulated parton-level events using \texttt{MadGraph5 v2.3.2.2} and
\texttt{MadEvent} \cite{Alwall:2014hca}, then passed those events to Pythia 6
\cite{Sjostrand:2006za} for showering and hadronization. Finally, we used
\texttt{Delphes 3} \cite{deFavereau:2013fsa} to perform a fast, parametrized
detector simulation, with the detector card devised by the FCC-hh working
group\footnote{\url{https://github.com/HEP-FCC/FCCSW/blob/master/Sim/SimDelphesInterface/data/FCChh_DelphesCard_Baseline_v01.tcl}}.
For the backgrounds, we allowed up to one additional jet in the final state, to
approximate NLO QCD effects, and performed MLM matching\footnote{Although
there are tools that support NLO generation \cite{Alwall:2014hca}, their NLO
generation mode does not support specifying decay chains, which we require in
order to generate enough MC events to be statistically representative of the
backgrounds at a 100 TeV $pp$ collider.} with the \texttt{xqcut}
parameter set to 40 GeV. The Higgsino pair production cross sections were
calculated using \texttt{Prospino2}~\cite{Beenakker:1999xh} at NLO.  
To decouple the Wino, we set its mass parameter $M_2$ to 3 TeV.

@@

Table \ref{tab:cc_cutflowtable} shows the cut efficiencies for a representative
signal benchmark point with $\mu =$ 1 TeV and $M_1$ = 25 GeV.   The
cross-sections of the backgrounds are obtained from MadEvent after performing
the MLM matching procedure. Their small size reflects the $m_{\ell\ell}$ and
$\slashed{E}_T$ cuts we imposed at the parton level. For all the benchmark
points, we require a minimum of three background events (at the MC level, i.e.
unweighted) left over after cuts\footnote{This is inspired by the upper
    one-sided limit for the mean of a Poisson variable given zero observed
    events, for a confidence level of 95\% (see Table 39.3 of
    \cite{PhysRevD.98.030001}), however, it does not reflect it exactly (since
    we required a minimum on unweighted rather than weighted events).  We
    observed that as we generated more unweighted background events, the
    discovery reach of our analysis showed a steady increase. This, combined
    with the fact that we were pushing the limits of our storage and computing
    capacity (we generated $\approx 10$ million $tt$ events, $\approx 26$
    million $tbW$  events and $100$K $bbWW$ events - all unweighted), lead us
    to believe that the projected reaches presented in this paper are
conservative ones. The ratio of the cross section of the generated events to
that expected at a 100 TeV $pp$ collider with 3 ab$^{-1}$ is $\approx 0.2$ for
$tt$, 0.05 for $tbW$, and $\approx0.46$ for $bbWW$.}. We can see that, after applying our
cuts, $tbW$ remains as the dominant backgrounds, followed by $tt$.  The values
of the razor variable cuts were chosen to maximize the significance,
$S/\sqrt{B}$ (calculated for an integrated luminosity of 3000 fb$^{-1}$), shown
in the last column.

@@

For each signal mass combination and each background process, we preselected
events that passed the lepton trigger, contained two SFOS leptons and two
$b$-tagged jets, and calculated a number of kinematic variables for each event.   We then
placed these kinematic variables in an array, with each row corresponding to an
event, and the columns corresponding to the kinematic variables. A mixture of
low-level and high-level kinematic variables was shown to have the greatest
effectiveness. The kinematic variables chosen were $m_{\ell\ell}$, $m_{bb}$,
$M_R$, $M_T^R$, $\slashed{E}_T,$ the total hadronic transverse energy $H_T$,
and the transverse momenta of individual final state leptons and $b$-quarks:
$p_T^{\ell_1}$, $p_T^{\ell_2}$, $p_T^{b_1}$, and $p_T^{b_2}$. We then divided
the generated events into training and test sets, with the training set
comprised of 75\% of the generated signal events and 30\% of the generated
background events. We require a lower portion of the background events for
training, in order to maintain adequate MC statistics to calculate significance
using the testing split, given the extremely large background cross sections at
a 100 TeV $pp$ collider. All the events are unweighted. We used the
\texttt{scikit-learn} package  \cite{Pedregosa2011} to implement our analysis.
A boosted decision tree classifier was trained with 1000 weak learners and a
learning rate of 0.025. The maximum tree depth is 3, and the minimum node
size is 1.

@@
